---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# sparsevinereg: High-dimensional sparse vine copula regression

<!-- badges: start -->
<!-- badges: end -->

An R package that selects variables from high-dimensional continuous data to make vine copula based (quantile) univariate predictions.

It depends on [vinereg](https://github.com/tnagler/vinereg) and [kde1d](https://github.com/tnagler/kde1d).

## Installation

You can install the development version from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("oezgesahin/sparsevinereg")
```


## Package overview

Below is an overview of some functions and features.

\begin{itemize}
\item \code{sparsevinereg}: selects most relevant variables from the given data to make predictions based on (parametric) vine copulas. Returns an object of class \code{sparsevinereg}. The class has the following methods:
\begin{itemize}
\item \code{print}: a brief overview of the model statistics.
\item \code{summary}: the final model's vine and margins.
\end{itemize}
\item \code{predict.sparsevinereg}: make predictions at a given quantile or mean level.
\end{itemize}

### Simulate data with 5 relevant and 5 irrelevant variables
```{r example2}
library(sparsevinereg)
set.seed(11)
x <- matrix(rnorm(5000), 500, 10)
y <- x[,1] -2*x[,2] + 3*x[,3] + 5*x[,4] - 4*x[,5]

# response is in the first column of the data
data <- data.frame(y = y, x = x)
```

###  fit a sparse vine copula based regression model 
```{r example3}
fit <- sparsevinereg(data)
```


###  fit a sparse vine copula based regression model based on partial correlations
```{r example4}
fit_ParCor <- sparsevinereg(data, varsel='ParCor')
```


###  print the main statistics of the model
```{r example5}
print(fit)
```


###  make predictions with the model at the quantile level 0.90
```{r example6}
upper_pred <- predict(fit, data, 0.90)
```

###  make predictions with the model at the mean 
```{r example7}
mean_pred <- predict(fit, data, NA)
```

## Contact
Please contact [ozgesahin-94@hotmail.com](mailto:ozgesahin-94@hotmail.com) if you have any questions.


## References
Sahin, O., and Czado, C. (2022). High-dimensional sparse vine copula regression with application to genomic prediction. arXiv preprint arXiv:2208.12383. [preprint](https://arxiv.org/pdf/2208.12383.pdf).
